---
title: "Support Vector Machines (SVMs)- Supervised Image Classification"
---

<div style="margin-bottom:40px;">
</div>

**Support Vector Machines (SVMs)** are supervised learning algorithms used mostly for classification problems. SVM models apply non-linear functions to select the best relationship between the response variable and predictors by introducing kernels functions  that map the independent variables to higher dimensional feature spaces . This approach typically leads to a better generalization of the chosen model on out-of-sample data . The principle behind an SVM classifier  algorithm is to separate data into different classes using a hyperplane. The goal in choosing a hyperplane is to maximize the distance from the hyperplane to the nearest data point of either class . These nearest data points are known as Support Vectors.


<div style="margin-bottom:42px;">
</div>

#### Load R packages 

```{r message=F, warning=F}
library(caret)        # machine laerning
library(kernlab)      # support vector machine
library(rgdal)        # spatial data processing
library(raster)       # raster processing
library(plyr)         # data manipulation 
library(dplyr)        # data manipulation 
library(RStoolbox)    # ploting spatial data 
library(RColorBrewer) # color
library(ggplot2)      # ploting
library(sp)           # spatial data
library(doParallel)   # Parallel processing
```


<div style="margin-bottom:20px;">
</div>

The data could be available for download from [here](https://www.dropbox.com/s/ybh1kr487wrhl8i/DATA_09.7z?dl=0).

<div style="margin-bottom:20px;">
</div>



```{r}
# Define data folder
dataFolder<-"F://Spatial_Data_Processing_and_Analysis_R//Data//DATA_09//"
```


#### Load data

```{r}
train.df<-read.csv(paste0(dataFolder,".\\Sentinel_2\\train_data.csv"), header = T)
test.df<-read.csv(paste0(dataFolder,".\\Sentinel_2\\test_data.csv"), header = T)
```

#### Start foreach to parallelize for model fitting

```{r}
mc <- makeCluster(detectCores())
registerDoParallel(mc)
```


#### Tunning prameters

```{r message=F, warning=F}
myControl <- trainControl(method="repeatedcv", 
                          number=3, 
                          repeats=2,
                          returnResamp='all', 
                          allowParallel=TRUE)
```

####  Train SVM model

We will use the train() function from the **caret** package with “method” parameter “**svmRadial**” (Radial Based Kernel based classification) wrapped  from the **Kernlab** package. 

```{r message=F, warning=F}
set.seed(849)
fit.svm <- train(as.factor(Landuse)~B2+B3+B4+B4+B6+B7+B8+B8A+B11+B12, 
                data=train.df,
                method = "svmRadial",
                metric= "Accuracy",
                preProc = c("center", "scale"), 
                trControl = myControl
                )
fit.svm 
```

#### Stop cluster

```{r}
stopCluster(mc)
```

#### Confusion Matrix - train data

```{r}
p1<-predict(fit.svm, train.df, type = "raw")
confusionMatrix(p1, train.df$Landuse)
```


#### Confusion Matrix - test data

```{r}
p2<-predict(fit.svm, test.df, type = "raw")
confusionMatrix(p2, test.df$Landuse)
```


#### Predition at grid location


```{r message=F, warning=F}
# read grid CSV file
grid.df<-read.csv(paste0(dataFolder,".\\Sentinel_2\\prediction_grid_data.csv"), header = T) 
# Preddict at grid location
p3<-as.data.frame(predict(fit.svm, grid.df, type = "raw"))
# Extract predicted landuse class
grid.df$Landuse<-p3$predict  
# Import lnaduse ID file 
ID<-read.csv(paste0(dataFolder,".\\Sentinel_2\\Landuse_ID.csv"), header=T)
# Join landuse ID
grid.new<-join(grid.df, ID, by="Landuse", type="inner") 
# Omit missing values
grid.new.na<-na.omit(grid.new)                                                            
```


#### Convert to raster 

```{r message=F, warning=F}
x<-SpatialPointsDataFrame(as.data.frame(grid.new.na)[, c("x", "y")], data = grid.new.na)
r <- rasterFromXYZ(as.data.frame(x)[, c("x", "y", "Class_ID")])
```

#### Plot Landuse Map: 

```{r,echo=TRUE,fig.align='center',fig.height=5, fig.width=6.5,message=F, warning=F}
# Color Palette
myPalette <- colorRampPalette(c("light grey","burlywood4", "forestgreen","light green", "dodgerblue"))
# Plot Map
LU<-spplot(r,"Class_ID", main="Supervised Image Classification: svmRadial" , 
      colorkey = list(space="right",tick.number=1,height=1, width=1.5,
              labels = list(at = seq(1,4.8,length=5),cex=1.0,
              lab = c("Road/parking/pavement" ,"Building", "Tree/buses", "Grass", "Water"))),
              col.regions=myPalette,cut=4)
LU
```


#### Write raster

```{r}
# writeRaster(r, filename = paste0(dataFolder,".\\Sentinel_2\\SVM_Landuse.tiff"), "GTiff", overwrite=T)
```




```{r}
rm(list = ls())
```
